OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-06-15 paper/paper.Rmd
Note gg hasn't been updated since 2013

When fitting to public databases, OTU quality dropped in closed reference mode
to different degrees depending on the database and dataset source, and no more
than half of query sequences were able to be fit to OTUs across any
dataset/database combination. 
This may reflect limitations of reference
databases, which are unlikely to contain sequences from rare and novel microbes.
This drop in quality was most notable with RDP, which contains only about 21,000
sequences compared to over 200,000 sequences in SILVA and Greengenes each at the
time of this writing.
**TODO:** note that greengenes is defunct now?!
We recommend that users who require an independent reference database opt for
large databases with good coverage of microbial diversity.
Since OptiClust performs faster than open reference OptiFit and creates higher
quality OTUs than closed reference OptiFit with the database strategy, we
recommend using OptiClust rather than fitting to a database whenever stable OTUs
are not required for the study at hand.

>>>>

When fitting to public databases, OTU quality dropped in closed reference mode
to different degrees depending on the database and dataset source, and no more
than half of query sequences were able to be fit to OTUs across any
dataset/database combination. 
This may reflect limitations of reference
databases, which are unlikely to contain sequences from rare and novel microbes.
This drop in quality was most notable with RDP, which contains only about 21,000
sequences compared to over 200,000 sequences in SILVA and Greengenes each.
Note that Greengenes has not been updated since 2013 at the time of this
writing, while SILVA is updated every one to two years with the most recent
release in 2019.
We recommend that users who require an independent reference database opt for
large databases with regular updates and good coverage of microbial diversity.
Since OptiClust performs faster than open reference OptiFit and creates higher
quality OTUs than closed reference OptiFit with the database strategy, we
recommend using OptiClust rather than fitting to a database whenever stable OTUs
are not required for the study at hand.
